#Reproducible Research Assignment 1#   

<br><br>   

####Loading and preprocessing the data
```{r,cache=TRUE}
setwd("C:/Users/NUS/Desktop/rdata")
activity<-read.csv("activity.csv")
activity$date<-as.Date(activity$date,"%Y-%m-%d")

```
<br><br>

####What is the mean total number of steps taken per day?
```{r,cache=TRUE}
## summing total number of steps by date using the aggregate command
aggregate(steps~date, data=activity,sum)->stepsbyday
with(stepsbyday,hist(steps,breaks=30,main="Histogram of Steps"))
```

```{r,cache=TRUE}
## mean and median of the total number of steps taken per day
meanstepsbyday<-as.integer(mean(stepsbyday$steps))
medianstepsbyday<-as.integer(median(stepsbyday$steps))
```

The mean and median of the total number of steps taken each day are `r meanstepsbyday` and `r medianstepsbyday` respectively.
<br><br>

####What is the average daily activity pattern?
```{r,cache=TRUE}
## obtaining the average number of steps in each interval across all days
byinterval<-aggregate(steps~interval,data=activity,mean)
byinterval$interval[which(byinterval$steps==max(byinterval$steps))]->answer
```
The 5-minute interval (averaged across all the days in the dataset), which contains the maximum number of steps is `r answer`.

```{r,cache=TRUE}
## Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
with (byinterval,plot(interval,steps,type="l", xlab="5-minute Intervals", ylab="Average Number of Steps Across All Days", main="Number of Steps for Each Interval, Averaged Across All Days"))

```
<br><br>



####Imputing missing values   

```{r,cache=TRUE}
sum(is.na(activity$steps))->missingvalue
```

The number of rows with missing values is `r missingvalue`.



```{r,cache=TRUE}
byinterval<-aggregate(steps~interval,data=activity,mean)
names(byinterval)<-c("interval","intervalmeansteps")
## merge the data frame containing the average number of steps for each interval across all days with the original activity dataset
activity1<-merge(activity,byinterval, by="interval")
suppressMessages(suppressWarnings(library(dplyr)))
activity1<-arrange(activity1, date) 
## replace the NA values with the average number of steps for that interval across all days
suppressWarnings(activity1$steps[is.na(activity1$steps)]<-activity1$intervalmeansteps) 
aggregate(steps~date, data=activity1,sum)->stepsbyday1  
with(stepsbyday1, hist(steps, breaks=30,main="Histogram of Steps (After Imputing Missing Values)")) 
```



```{r,cache=TRUE}
meanstepsbyday1<-as.integer(mean(stepsbyday1$steps))
medianstepsbyday1<-as.integer(median(stepsbyday1$steps))

```
After imputing the missing values, the mean and median of the total number of steps taken each day are `r meanstepsbyday1` and `r medianstepsbyday1` respectively.
<br><br>



####Are there differences in activity patterns between weekdays and weekends?

```{r,cache=TRUE}
activity1$date<-as.Date(activity1$date,"%Y-%m-%d")
## Use the weekdays function from the lubridate package to determine the days for ## each date.
suppressMessages(library(lubridate))
activity1$days<-weekdays(activity1$date)
weekday<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
## generate a new variable daysfactor which takes on the value of "weekday"" from ## Monday to Fridays and take on the value of "weekend" for Saturdays and Sundays
activity1$daysfactor[activity1$days %in% weekday]<-"weekday"
weekend<-c("Saturday","Sunday")
activity1$daysfactor[activity1$days %in% weekend]<-"weekend"
activity1$daysfactor<-as.factor(activity1$daysfactor)
## subset the data into two dataframes based on data depending on whether each row corresponds to a weekday or weekend
weekend_activity1<- subset(activity1, daysfactor=="weekend")
weekday_activity1<- subset(activity1, daysfactor=="weekday")
## compute the average number of steps in each interval across all weekdays and ## all weekends
byinterval_weekend<-aggregate(steps~interval,data=weekend_activity1,mean)
byinterval_weekday<-aggregate(steps~interval,data=weekday_activity1,mean)

## generate a dataframe with the average number of steps in each interval across ## all days, with a factor variable that indicates whether each row corresponds to ## a weekend or weekday.
byinterval_weekend$factor<-"weekend"
byinterval_weekday$factor<-"weekday"
compare<-rbind(byinterval_weekday,byinterval_weekend)

## Draw a panel plot containing two time series plots, 
library(ggplot2)
g<-ggplot(data=compare, aes(interval,steps,color=factor))
g<-g+geom_line()+xlab("Interval")+ylab("Number of Steps") + ggtitle("Differences in Activity Patterns on Weekdays and Weekends")
g<-g+facet_grid(factor~.)+theme_bw()
print(g)
```



